Revisiting the hydrodynamic wave modulation from linear wave theory

### 2024-08-16 paper/paper.tex
Update abstract; captions etc.

\begin{keypoints}
\item We revisit analytical and numerical solutions for hydrodynamic modulation of short waves by longer waves.
\item We present a higher-order analytical solution based on linear wave theory.
\item In contrast to an earlier study, we find stable numerical solutions for modulation by realistic long-wave groups.
\end{keypoints}

\begin{abstract}
Hydrodynamic modulation of short ocean surface waves by longer ambient waves is
a well-known ocean surface process that pertains to remote sensing, the
interpretation of in situ wave measurements, and numerical wave forecasting.
In this paper we revisit the early analytical work by \citeA{longuet1960changes},
as well as the more recent numerical simulations by \citeA{peureux2021unsteady},
to re-evaluate the changes in short-wave properties due to the presence of longer waves.
We present two distinct improvements to the current state of the art:
First, we derive a higher-order, non-linear solution for the short-wave wavenumber
and action density as function of the long-wave phase and steepness.
This solution suggests an increase in short-wave wavenumber by 18 and 38\% for
long-wave steepenesses of 0.1 and 0.2, respectively, compared to the original
linear solution.
It is also in close agreement with the full numerical solutions up to the
long-wave steepness of 0.25.
Second, we find that the full numerical solutions, which were previously found
to be unstable for homogeneous initial conditions, are in fact stable for more
realistic long-wave groups.
The key physical interpretation from these results is that the short waves can
indefinitely grow and steepen only in the presence of inifinite long-wave trains,
an unlikely occurence in the ocean.
\end{abstract}

\section*{Plain Language Summary}
The ocean surface is typically covered by waves of many wavelengths.
Longer waves bunch up and dilate the shorter waves, and this process is
important to consider for both the measurements and the forecasting of ocean
waves.
Although there have been analytical and numerical solutions for how the short
waves should change in the presence of longer waves, they have been simplified
and incomplete.
This paper revisits and extends these solutions to propose a more accurate
description of how the short waves should change in the presence of longer
waves.

@@

\begin{table}[h]
\label{table:modulation_literature_summary}
\caption{
  Summary of prior \cite{longuet1960changes,longuet1987propagation,zhang1990evolution} and present (this paper)
  approaches to calculating the modulation of short waves by long waves.
}
\begin{tabular}{cccccc}
\hline
Reference   & Governing & Gravity & Nonlinear  & Nonlinear   & Solution \\
            & equations & varies  & long waves & short waves &          \\
\hline
\hline
L-HS (1960) & Velocity potential       & No  & No  & No  & Analytical \\
\hline
L-H (1987)  & Velocity potential       & Yes & Yes & No  & Numerical \\
\hline
ZM (1990)   & Nonlinear Schr\" odinger & Yes & Yes & Yes & Numerical \\
            & equation                 &     &     &     &           \\
\hline
This paper  & Crest and action         & Yes & Yes & No  & Analytical \\
            & conservation             &     &     &     & and numerical \\
\hline
\end{tabular}
\end{table}

@@

The second important relation is the conservation of wave crests:

@@

\begin{figure}[ht]
\label{fig:analytical_solutions_ak0.1}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.1.png}
\caption{
}
\end{figure}

@@

\begin{figure}[ht]
\label{fig:analytical_solutions_ak0.2}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.2.png}
\caption{Same a Fig. \ref{fig:analytical_solutions_ak0.1} but for $\varepsilon_L = 0.2$.}
\end{figure}

\begin{figure}[ht]
\label{fig:analytical_solutions_ak0.4}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.4.png}
\caption{Same a Fig. \ref{fig:analytical_solutions_ak0.1} but for $\varepsilon_L = 0.4$.}
\end{figure}

@@

\begin{figure}[h]
  \label{fig:modulation_3panel_ramp}
  \centering
  \includegraphics[width=\textwidth]{figures/fig_modulation_ramp_timeseries.png}
  \caption{
    Maximum short-wave steepness modulation as function of time in case of
    infinite long-wave train case (blue) and the same case but with a linear
    ramp during the initial five long-wave periods (orange). The dashed black
    line corresponds to the short-wave steepness of 0.4 at which most waves are
    expected to break.
  }
  \end{figure}

@@

\acknowledgments
Thanks to Peisen Tan, Nathan Laxague, and Fabrice Ardhuin for providing helpful
feedback on the early drafts of this work.

>>>>

\begin{keypoints}
\item We revisit analytical and numerical solutions for hydrodynamic modulation of short waves by longer waves.
\item We present a higher-order analytical solution based on linear wave theory.
\item In contrast to an earlier study, we find stable numerical solutions for modulation by long-wave groups.
\end{keypoints}

\begin{abstract}
Hydrodynamic modulation of short ocean surface waves by longer ambient waves is
a well-known ocean surface process that affects remote sensing, the
interpretation of \textit{in situ} wave measurements, and numerical wave
forecasting.
In this paper, we revisit the linear wave theory and derive higher-order
solutions for the change of short-wave wavenumber, action density, and
gravitational acceleration due to the presence of longer waves.
We validate the analytical solutions with numerical simulations of the full wave
crest and action conservation equations.
The nonlinear analytical solutions of short-wave wavenumber, amplitude, and
steepness modulation significantly deviate from the linear analytical solutions
of \citeA{longuet1960changes}, and are similar to the nonlinear numerical
solutions by \citeA{zhang1990evolution}.
We also discuss the result of \citeA{peureux2021unsteady} who found through
numerical simulations that the short-wave modulation grows unsteadily with
each long-wave phase.
Here we show that this unsteady growth only occurs for homogeneous initial
conditions and does not occur for more realistic long-wave groups.
Finally, the combined analytical and numerical framework allows us to quantify
the contribution of various terms in the full wave conservation equations.
The new analytical solutions are a good approximation of the full numerical
solution in long-wave steepness up to $\approx 0.2$.
\end{abstract}

\section*{Plain Language Summary}
The ocean surface is typically covered by waves of many wavelengths.
Longer waves bunch up and dilate the shorter waves, and this process is
important to consider for both the measurements and the forecasting of ocean
waves.
Although there have been analytical and numerical solutions for how the short
waves should change in the presence of longer waves, they have either been
simplistic or 
This paper revisits and extends these solutions to propose a more accurate
description of how the short waves should change in the presence of longer
waves.

@@

\begin{table}[h]
\label{table:modulation_literature_summary}
\caption{
  Summary of prior
  \cite{longuet1960changes,phillips1981dispersion,longuet1987propagation,zhang1990evolution,peureux2021unsteady}
  and present (this paper) approaches to calculating the modulation of short
  waves by long waves.
  *: long 3rd order Stokes wave is considered only in numerical simulations.
}
\begin{tabular}{cccccc}
\hline
Reference   & Governing & Gravity & Nonlinear  & Nonlinear   & Solution \\
            & equations & varies  & long waves & short waves &          \\
\hline
\hline
L-HS 1960   & Velocity potential       & No  & No  & No  & Analytical \\
\hline
P 1981      & Crest and energy         & Yes & Yes & No  & Numerical \\
            & conservation             &     &     &     &           \\
\hline
L-H 1987    & Velocity potential       & Yes & Yes & No  & Numerical \\
\hline
ZM 1990     & Nonlinear Schr\" odinger & Yes & Yes & Yes & Numerical \\
            & equation                 &     &     &     &           \\
\hline
PAG 2021    & Crest and action         & Yes & No & No  & Numerical \\
            & conservation             &     &     &     &           \\
\hline
This paper  & Crest and action         & Yes & Yes* & No  & Analytical \\
            & conservation             &     &      &     & and numerical \\
\hline
\end{tabular}
\end{table}

@@

The second important relation is the conservation of wave crests
\cite{phillips1981dispersion}:

@@

\begin{figure}[ht]
\label{fig:analytical_solutions_ak0.1}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.1.png}
\caption{
  Modulation of short-wave (a) wavenumber, (b) gravitational acceleration,
  (c) amplitude, (d) steepness, (e) intrinsic frequency, and (d) intrinsic phase
  speed as function of long-wave phase for $\varepsilon_L = 0.1$, based on
  analytical solutions by \citeA{longuet1960changes} (L-HS 1960, black)
  and this paper (blue).
  Long-wave crest and trough are located at $\psi = 0$ and $\psi = \pi$,
  respectively.
}
\end{figure}

@@

\begin{figure}[ht]
\label{fig:analytical_solutions_ak0.2}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.2.png}
\caption{As Fig. \ref{fig:analytical_solutions_ak0.1} but for $\varepsilon_L = 0.2$.}
\end{figure}

\begin{figure}[ht]
\label{fig:analytical_solutions_ak0.4}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_analytical_solutions_ak0.4.png}
\caption{As Fig. \ref{fig:analytical_solutions_ak0.1} but for $\varepsilon_L = 0.4$.}
\end{figure}

@@

\begin{figure}[h]
  \label{fig:unsteady_growth_timeseries}
  \centering
  \includegraphics[width=0.6\textwidth]{figures/fig_modulation_ramp_timeseries.png}
  \caption{
    Maximum short-wave steepness modulation as function of time in case of
    infinite long-wave train case (blue) and the same case but with a linear
    ramp during the initial five long-wave periods (orange). The dashed black
    line corresponds to the short-wave steepness of 0.4 at which most waves are
    expected to break.
  }
  \end{figure}

@@

\acknowledgments
Thanks to Nathan Laxague, Fabrice Ardhuin, Nick Pizzo, and Peisen Tan for the
helpful discussions.
